% ============================================================
% Consciousness Coherence in the Generative-Collapse Kernel:
% Seven Tier-2 Theorems from a Corrected Closure
%
% A rigorous GCD rewrite of:
%   Jackson (2026a) "The Year the Earth Stood Still"
%   Jackson (2026b) "The Kannsas Factor κ_i"
%   Jackson (2026c) "UMCP/GCD Tier-0 Adapter Specification v1.1"
%
% Compile: pdflatex → bibtex → pdflatex → pdflatex
% ============================================================

\documentclass[
  aps,
  prd,
  twocolumn,
  superscriptaddress,
  nofootinbib,
  floatfix
]{revtex4-2}

% ── packages ──
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{graphicx}
\usepackage{xcolor}
\definecolor{linkblue}{HTML}{337799}
\usepackage[colorlinks=true,linkcolor=linkblue,citecolor=linkblue,urlcolor=linkblue]{hyperref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xfrac}
\usepackage{enumitem}
\usepackage{array}
\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}

% ── theorem environments ──
\newtheorem*{axiom}{The Axiom}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{identity}[theorem]{Identity}
\newtheorem{remark}{Remark}

% ── UMCP macros ──
\newcommand{\tR}{\tau_{\!R}}
\newcommand{\tRS}{\tau_{\!R}^{*}}
\newcommand{\Gam}{\Gamma}
\newcommand{\eps}{\varepsilon}
\newcommand{\IR}{\infty_{\mathrm{rec}}}
\newcommand{\tolseam}{\mathrm{tol}_{\mathrm{seam}}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Res}{\mathrm{Res}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\INF}{\texttt{INF\_REC}}
\newcommand{\regime}[1]{\textrm{\upshape\scshape#1}}
\newcommand{\conform}{\regime{conformant}}
\newcommand{\nonconform}{\regime{nonconformant}}
\newcommand{\tvec}{\mathbf{c}}
\newcommand{\wvec}{\mathbf{w}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\gF}{g_{\!F}}
\newcommand{\RunID}{\mathrm{RunID}}
\newcommand{\xiJ}{\xi_J}

\begin{document}

% ── title ──
\title{Consciousness Coherence in the Generative-Collapse Kernel:\\
Seven Tier-2 Theorems from a Corrected Closure}

\author{Clement Paulus}
\email{clementpaulus9@gmail.com}
\affiliation{UMCP Project}

\date{\today}

% ── abstract ──
\begin{abstract}
Jackson~(2026) proposes $\xiJ = 7.2$ ($= 432/60$) as a ``universal
consciousness coherence constant,'' claiming it unifies $\varphi^2 e$,
base-60 arithmetic, 432\,Hz tuning, and recursive self-reference into a
``golden harmonic web.''  We recast these claims as a Tier-2 closure
within the Generative Collapse Dynamics (GCD) kernel, correcting
arithmetic errors (the reported $e^{i\xiJ}+1 = 1.673+0.739i$ is actually
$1.608+0.794i$), resolving a symbol collision ($\xiJ$, not $\kappa$),
and testing every claim computationally.  Seven theorems emerge from 20
consciousness-candidate systems through an 8-channel trace vector:
(T-CC-1)~$\xiJ$ proximity does not predict composite integrity
($\rho = +0.25$, $p = 0.28$);
(T-CC-2)~recursive depth anti-correlates with return fidelity
($\rho = -0.10$);
(T-CC-3)~no system reaches Stable regime (0/20);
(T-CC-4)~one dead channel kills coherence (geometric slaughter);
(T-CC-5)~432\,Hz and 440\,Hz are kernel-equivalent
($|\Delta F| = 0.013$);
(T-CC-6)~mathematical identity outperforms all harmonic systems; and
(T-CC-7)~the heterogeneity gap inverts Jackson's hierarchy
(Neural $\Delta = 0.015$, Harmonic $\Delta = 0.114$).
All 7/7 theorems are proven with 38/38 sub-tests.  The corrected
thesis---``do harmonic-ratio systems exhibit measurable coherence
patterns?''---is a legitimate Tier-2 question whose answer is: no,
the harmonic ratio $\xiJ$ is not structurally privileged; coherence
is predicted by channel homogeneity ($\Delta = F - \IC$), not by
proximity to any fixed constant.
\end{abstract}

\maketitle

% ════════════════════════════════════════════════════════════
\section{Introduction}
% ════════════════════════════════════════════════════════════

Jackson~\cite{jackson2026earth,jackson2026curved} proposes that a single
constant $\xiJ = 7.2$---equal to $432/60$ in the Babylonian base-60
system and approximately $\varphi^2 e \approx 7.117$---governs
consciousness coherence across biological, physical, and mathematical
domains.  The thesis is presented through the REMIS framework (Recursive
Emergent Metacognitive Intelligence Study) and extended with OBE (out-of-body
experience) mechanics, 432\,Hz acoustic tuning claims, and a ``golden
harmonic web'' connecting $\varphi$, $e$, $\pi$, and the speed of light.

The claims contain several arithmetic errors, a symbol collision with GCD's
reserved $\kappa$, unit-dependent artifacts, and a Spine that cannot be
evaluated (no frozen contract, no integrity ledger, no derived stance).
Our prior assessment classified the thesis as a \emph{gesture}---the
questions are legitimate but none of the claims demonstrates return.

This paper takes a different approach: rather than dismissing the claims as
gestures, we \emph{correct} each one and \emph{ground} it in the GCD
kernel.  Every Jackson claim is mapped to a falsifiable computation on
20~consciousness-candidate systems through an 8-channel trace vector.
Seven theorems emerge, all proven, revealing what the kernel actually says
about consciousness coherence---and what of Jackson's intuitions survives
the correction.

% ════════════════════════════════════════════════════════════
\section{Symbol Resolution}
\label{sec:symbols}
% ════════════════════════════════════════════════════════════

Jackson uses $\kappa$ for his constant.  In GCD, $\kappa$ is a Tier-1
reserved symbol~\cite{paulus2025umcp}:
\beq
  \kappa = \sum_i w_i \ln(c_{i,\eps}), \qquad \kappa \leq 0
\eeq
where $c_{i,\eps} = \max(c_i, \eps)$ is the $\eps$-clamped channel value.
Jackson's $\kappa = 7.2$ is positive and fixed; GCD's $\kappa$ is variable
and non-positive.  The ranges do not overlap.  Worse,
$\IC = \exp(\kappa) = \exp(7.2) \approx 1339$, violating $\IC \in (0,1]$.

\begin{remark}[Symbol Resolution]
We rename Jackson's constant to $\xiJ$ (``xi-Jackson'') throughout.
$\xiJ = 432/60 = 7.2$ is a Tier-2 domain parameter, not a frozen constant.
\end{remark}

% ════════════════════════════════════════════════════════════
\section{Arithmetic Corrections}
\label{sec:corrections}
% ════════════════════════════════════════════════════════════

Table~\ref{tab:corrections} lists every numerical claim from
Jackson~\cite{jackson2026curved} with the corrected value computed
from first principles.

\begin{table}[h]
\caption{Corrected Jackson arithmetic.  Every claim is re-derived from
the defining equations.  Bold entries differ significantly from Jackson's
reported values.}
\label{tab:corrections}
\begin{ruledtabular}
\begin{tabular}{lcc}
\textrm{Claim} & \textrm{Jackson} & \textrm{Correct} \\
\colrule
$\varphi^2 e$                    & ${\approx}\,7.2$   & $\mathbf{7.117}$ (1.16\%) \\
$\mathrm{Re}(e^{i\xiJ}+1)$      & $1.673$          & $\mathbf{1.608}$              \\
$\mathrm{Im}(e^{i\xiJ}+1)$      & $0.739$          & $\mathbf{0.794}$              \\
Excess angle $(\xiJ - 2\pi)$     & $47.68^\circ$    & $\mathbf{52.53}^\circ$        \\
Cycles for $360^\circ$           & ${\sim}\,7.5$     & $\mathbf{6.85}$               \\
$c = \alpha^{-1}{\cdot}432{\cdot}\pi$ & ${\approx}\,c$ (km/s) & unit artifact (38\% off) \\
$432/\xiJ$                       & $60$ (exact)     & $60$ (trivially true)         \\
$\varphi^2/\xiJ$                 & $0.364$          & $0.364$ (ratio)               \\
\end{tabular}
\end{ruledtabular}
\end{table}

The pattern is consistent: claims that appear exact under inspection turn
out to be either near-misses (1.16\%), arithmetic errors (Euler rotation),
trivial identities ($432/60$), or unit artifacts (speed of light).

% ════════════════════════════════════════════════════════════
\section{The Corrected Thesis}
\label{sec:thesis}
% ════════════════════════════════════════════════════════════

Jackson asks: \emph{Is there a special constant that unifies consciousness
across biological, mathematical, and physical systems?}

The corrected, falsifiable version: \emph{Do systems exhibiting harmonic
structure near $\xiJ = 7.2$ show measurably higher composite integrity
$\IC$ than systems without such structure?}

This question can be answered by the GCD kernel.  We map 20
consciousness-candidate systems into 8 channels and test whether
$\xiJ$~proximity predicts kernel coherence.

% ════════════════════════════════════════════════════════════
\section{Channel Construction}
\label{sec:channels}
% ════════════════════════════════════════════════════════════

Each system maps to an 8-channel trace vector
$\tvec \in [\eps, 1-\eps]^8$ with equal weights
$\wvec = (1/8, \ldots, 1/8)$:

\begin{enumerate}
  \item \textbf{harmonic\_ratio}: proximity of system frequency ratios
        to $\xiJ = 7.2$; $[0, 1]$
  \item \textbf{recursive\_depth}: demonstrated self-referential
        structural depth; $[0, 1]$
  \item \textbf{return\_fidelity}: $\tR$ finiteness---does the system
        re-enter prior states?; $[0, 1]$
  \item \textbf{spectral\_coherence}: power spectrum concentration
        vs.\ uniform noise; $[0, 1]$
  \item \textbf{phase\_stability}: phase-locking across coupled
        oscillators; $[0, 1]$
  \item \textbf{information\_density}: bits per symbol in the system's
        encoding; $[0, 1]$
  \item \textbf{temporal\_persistence}: duration of coherent state
        maintenance; $[0, 1]$
  \item \textbf{cross\_scale\_coupling}: coherence preserved across
        spatial/temporal scales; $[0, 1]$
\end{enumerate}

The 20 systems span 5~categories: Neural (6), Harmonic (3), Physical~(5),
Mathematical (3), and Recursive (3).  Each system profile is documented
in the closure module with normalization conventions.

% ════════════════════════════════════════════════════════════
\section{Seven Theorems}
\label{sec:theorems}
% ════════════════════════════════════════════════════════════

\subsection{T-CC-1: Harmonic Non-Privilege}

\begin{theorem}[Harmonic Non-Privilege]
The Spearman rank correlation between $\xiJ$~proximity (harmonic\_ratio
channel) and composite integrity $\IC$ across 20~systems is
$\rho = +0.25$, $p = 0.28$---not statistically significant.
The near-miss $\varphi^2 e = 7.1166 \neq 7.2$ has a 1.16\% residual,
exceeding the machine-precision threshold for structural identities.
\end{theorem}

\emph{Jackson's claim}: $\kappa = \varphi^2 e \approx 7.2$ is a universal
consciousness coherence constant.

\emph{Correction}: $\xiJ = 432/60$ is a harmonic ratio in base-60
arithmetic.  $\varphi^2 e = 7.1166$ is a near-miss, not an identity.
The constant is not privileged by the data.  Compare: the GCD duality
identity $F + \omega = 1$ holds to $< 10^{-16}$ residual across all
systems---\emph{that} is what an identity looks like.

\emph{Key numbers}: $\rho = +0.25$, $p = 0.28$; Jackson's own system
is in Collapse regime with $F = 0.455$, $\IC = 0.400$.

\subsection{T-CC-2: Recursion-Return Dissociation}

\begin{theorem}[Recursion-Return Dissociation]
The Spearman rank correlation between recursive\_depth (channel~2) and
return\_fidelity (channel~3) is $\rho = -0.10$, $p = 0.68$.  Systems
exhibiting high recursive depth ($> 0.60$) but low return fidelity
($< 0.50$) exist and occupy Collapse regime (Recursive Gestures).
\end{theorem}

\emph{Jackson's claim}: REMIS posits that recursive self-reference
produces consciousness.

\emph{Correction}: Recursion is a \emph{channel}, not a \emph{gate}.
What matters in GCD is \textbf{return} ($\tR$ finite), not recursion
depth.  A system that recurses but never returns has $\tR = \infty_{\rm rec}$
and receives zero budget credit.  The LLM system (recursive\_depth $= 0.75$,
return\_fidelity $= 0.45$) is the poster child: a Recursive Gesture in
Collapse regime.  Meanwhile, the cardiac rhythm (recursive\_depth $= 0.20$,
return\_fidelity $= 0.95$) returns reliably without recursion.

\subsection{T-CC-3: Universal Instability}

\begin{theorem}[Universal Instability of Coherence]
No consciousness-candidate system ($0/20$) reaches the Stable regime.
The distribution is Watch~$= 5$, Collapse~$= 15$.  Even the highest-$F$
system (Euler identity, $F = 0.896$) remains in Watch due to the
conjunctive Stable gate ($\omega < 0.038 \land F > 0.90 \land
S < 0.15 \land C < 0.14$).
\end{theorem}

\emph{Jackson's claim}: The consciousness constant implies stable,
self-sustaining coherent states.

\emph{Correction}: Coherence is structurally unstable.  All 20~systems
live in the $87.5\%$ of Fisher space outside Stable.  This is not a
failure---it is Axiom-0 in action: ``only what returns is real.''
Consciousness persists through return from collapse, not through
avoidance of collapse.  \emph{Ruptura est fons constantiae}---dissolution
is the source of constancy.

\subsection{T-CC-4: Geometric Slaughter}

\begin{theorem}[Geometric Slaughter in Coherence]
The ratio $\IC/F$ (multiplicative vs.\ arithmetic coherence)
anti-correlates with channel variance: $\rho = -0.91$.  One near-zero
channel kills $\IC$ regardless of how strong the other channels are.
The Euler identity has $\IC/F = 0.993$ (most homogeneous); the 432\,Hz
system has $\IC/F = 0.846$ (information\_density $= 0.15$ kills it).
\end{theorem}

\emph{Jackson's claim}: Four ``independent'' derivations ($\varphi^2 e$,
$432/60$, speed of light, Euler rotation) reinforce $\kappa = 7.2$
multiplicatively.

\emph{Correction}: Four channels of one claim.  The weakest
channel---arithmetic errors ($e^{i\xiJ}+1$ wrong by $4\%$), unit
artifacts (speed of light in km/s only), and undemonstrated return
($\tR = \infty_{\rm rec}$)---kills the composite integrity through
geometric slaughter.  $\IC$ is a geometric mean: one dead channel $\to$
$\IC \to \eps$, regardless of the others.  Jackson's thesis has
$\IC/F = 0.879$, dragged down by cross\_scale\_coupling~$= 0.20$ and
return\_fidelity~$= 0.30$.

\subsection{T-CC-5: Tuning Invariance}

\begin{theorem}[Tuning Invariance: 432\,Hz $\equiv$ 440\,Hz]
The kernel signatures of 432\,Hz and 440\,Hz tuning systems are
indistinguishable: $|\Delta F| = 0.013$, $|\Delta\IC| = 0.008$,
same regime (Collapse), same weakest channel (information\_density),
$|\Delta\Delta| = 0.005$.
\end{theorem}

\emph{Jackson's claim}: 432\,Hz tuning is privileged because
$432/60 = \xiJ = 7.2$.

\emph{Correction}: The 8\,Hz difference between 432\,Hz and 440\,Hz is
not structurally significant.  Both systems occupy Collapse regime
with nearly identical $F$, $\IC$, $\Delta$, and the same weakest channel.
The number 432 is privileged only in base-60 arithmetic ($432 = 7.2
\times 60$), which is a property of the number system, not of acoustics.

\subsection{T-CC-6: Mathematical Supremacy}

\begin{theorem}[Mathematical Supremacy]
The Mathematical category has the highest mean $F$ ($0.690$) and highest
mean $\IC$ ($0.666$) of all five categories.  The Harmonic category has
the highest mean $\Delta$ ($0.114$), making it the most fragile.  The
Euler identity ($F = 0.896$, $\IC = 0.890$) outperforms every harmonic
system in both $F$ and $\IC$.
\end{theorem}

\emph{Jackson's claim}: Harmonic ratios (especially $\xiJ$) are the key
to consciousness coherence.

\emph{Correction}: Mathematical identity---exact, self-consistent
relations like $e^{i\pi}+1=0$---produces higher kernel coherence than
harmonic ratios.  Harmonic systems are the most \emph{fragile} category
(highest~$\Delta$) because they are specialists: high spectral\_coherence
($0.95$) but near-zero information\_density ($0.15$).  Geometric slaughter
penalizes this specialization.  Channel homogeneity, not harmonic proximity,
predicts integrity.

\subsection{T-CC-7: Heterogeneity Gap Ordering}

\begin{theorem}[Heterogeneity Gap Inverts Jackson's Hierarchy]
The category ordering by heterogeneity gap $\Delta = F - \IC$ is:
\beq
  \underbrace{\text{Neur.}}_{0.015} <
  \underbrace{\text{Math.}}_{0.024} <
  \underbrace{\text{Recur.}}_{0.076} <
  \underbrace{\text{Phys.}}_{0.111} <
  \underbrace{\text{Harm.}}_{0.114}
\eeq
This inverts Jackson's hierarchy: neural (consciousness-bearing) systems
have the lowest $\Delta$ (most internally coherent), while harmonic
systems (Jackson's preferred carrier) have the highest $\Delta$ (most
fragile).  Jackson's proposed threshold $\varphi^2/\xiJ = 0.364$ exceeds
the maximum $\Delta$ in the dataset ($0.166$) and is therefore vacuous.
The heterogeneity gap is more predictive of $\IC$ than $\xiJ$~proximity
($|\rho_{\Delta}| > |\rho_{\xiJ}|$).
\end{theorem}

\emph{Jackson's claim}: $\varphi^2/\kappa = 0.364$ is a consciousness
threshold.

\emph{Correction}: $\varphi^2/\xiJ = 0.364$ is a ratio of two Tier-2
parameters (neither frozen nor seam-derived).  It exceeds every $\Delta$
in the 20-system dataset---it is not a threshold.  The native GCD metric
$\Delta = F - \IC$ inverts Jackson's hierarchy: biology beats harmonics.
Neural systems need all 8~channels to work together, producing low~$\Delta$.
Harmonic systems sacrifice information density for spectral purity,
producing high~$\Delta$.

% ════════════════════════════════════════════════════════════
\section{What Survives}
\label{sec:survives}
% ════════════════════════════════════════════════════════════

Despite the corrections, four elements of Jackson's work make contact
with GCD structure:

\begin{enumerate}
  \item \textbf{Recursive self-reference as a measurable channel.}
        Jackson's REMIS thesis, stripped of numerological overlay, asks
        whether recursion correlates with coherence.  The answer from
        T-CC-2 is ``no---recursion does not predict return''---but the
        \emph{question} maps directly to the GCD kernel.  A refined
        version (``under what conditions does recursion produce finite
        $\tR$?'') is a legitimate open question.

  \item \textbf{Return through collapse as a structural requirement.}
        Jackson's Addendum~A.3 (OBE mechanics) describes consciousness
        state transitions as crossings that must ``return.''  This is
        Axiom-0's central claim, and T-CC-3 confirms it: coherence
        persists through return from collapse, not through stability.
        The concept needs a frozen contract and measured trace vectors
        to become a weld.

  \item \textbf{The heterogeneity gap as imperfection.}
        Jackson's remark that consciousness requires ``perfect
        imperfection'' maps to $\Delta = F - \IC$.  T-CC-7 shows that
        neural systems have the lowest $\Delta$ (most balanced channels),
        while harmonic systems have the highest (most specialist).  The
        gap is real and measurable; it just inverts Jackson's ordering.

  \item \textbf{Cross-domain kernel comparison.}
        Jackson's aspiration to compare consciousness across domains
        (biological, mathematical, physical) is precisely what the
        GCD kernel does.  The five categories in the closure produce
        comparable $F$, $\IC$, and $\Delta$ values on a common scale.
        This is the Rosetta adapter in action.
\end{enumerate}

% ════════════════════════════════════════════════════════════
\section{Summary of Results}
\label{sec:results}
% ════════════════════════════════════════════════════════════

\begin{table}[h]
\caption{Seven theorems: all 7/7 proven, 38/38 sub-tests.}
\label{tab:summary}
\begin{ruledtabular}
\begin{tabular}{lccc}
\textrm{Theorem} & \textrm{Tests} & \textrm{Pass} & \textrm{Verdict} \\
\colrule
T-CC-1 Harmonic Non-Priv.       & 5  & 5  & PROVEN \\
T-CC-2 Recur.-Return Dissoc.    & 5  & 5  & PROVEN \\
T-CC-3 Universal Instability     & 5  & 5  & PROVEN \\
T-CC-4 Geometric Slaughter       & 6  & 6  & PROVEN \\
T-CC-5 Tuning Invariance         & 6  & 6  & PROVEN \\
T-CC-6 Mathematical Supremacy    & 6  & 6  & PROVEN \\
T-CC-7 Het.\ Gap Ordering        & 5  & 5  & PROVEN \\
\colrule
\textbf{Total} & \textbf{38} & \textbf{38} & \textbf{7/7} \\
\end{tabular}
\end{ruledtabular}
\end{table}

Table~\ref{tab:kernel} shows the kernel output for all 20~systems.

\begin{table*}[t]
\caption{Kernel output for 20 consciousness systems.  $F$ = fidelity,
$\IC$ = composite integrity, $\Delta = F - \IC$ = heterogeneity gap,
$\omega$ = drift, $\kappa$ = log-integrity (GCD, not Jackson's $\xiJ$).
All systems satisfy $F + \omega = 1$, $\IC \leq F$, $\IC = \exp(\kappa)$.}
\label{tab:kernel}
\begin{ruledtabular}
\begin{tabular}{llcccccl}
\textrm{System} & \textrm{Cat.} & $F$ & $\IC$ & $\Delta$ & $\omega$ & $\kappa$ & \textrm{Regime} \\
\colrule
Human waking         & Neural   & 0.731 & 0.721 & 0.011 & 0.269 & $-0.328$ & Watch    \\
Human sleep (REM)    & Neural   & 0.488 & 0.473 & 0.015 & 0.513 & $-0.749$ & Collapse \\
Human meditation     & Neural   & 0.738 & 0.720 & 0.017 & 0.263 & $-0.328$ & Watch    \\
Cetacean             & Neural   & 0.575 & 0.564 & 0.011 & 0.425 & $-0.573$ & Collapse \\
Corvid               & Neural   & 0.475 & 0.456 & 0.019 & 0.525 & $-0.786$ & Collapse \\
Octopus              & Neural   & 0.300 & 0.284 & 0.016 & 0.700 & $-1.258$ & Collapse \\
432\,Hz              & Harmonic & 0.669 & 0.566 & 0.103 & 0.331 & $-0.569$ & Collapse \\
440\,Hz              & Harmonic & 0.656 & 0.558 & 0.098 & 0.344 & $-0.583$ & Collapse \\
Coral Castle         & Harmonic & 0.471 & 0.329 & 0.142 & 0.529 & $-1.112$ & Collapse \\
Circadian rhythm     & Physical & 0.646 & 0.573 & 0.073 & 0.354 & $-0.557$ & Collapse \\
Cardiac rhythm       & Physical & 0.625 & 0.530 & 0.095 & 0.375 & $-0.635$ & Collapse \\
Laser coherent light & Physical & 0.574 & 0.408 & 0.166 & 0.426 & $-0.897$ & Collapse \\
Euler identity       & Math     & 0.896 & 0.890 & 0.006 & 0.104 & $-0.116$ & Watch    \\
Jackson $\xiJ$       & Math     & 0.455 & 0.400 & 0.055 & 0.545 & $-0.916$ & Collapse \\
G\"odel self-ref      & Recursive & 0.716 & 0.610 & 0.107 & 0.284 & $-0.495$ & Watch    \\
CA Rule~110          & Recursive & 0.575 & 0.516 & 0.059 & 0.425 & $-0.662$ & Collapse \\
LLM dialogue         & Recursive & 0.444 & 0.382 & 0.062 & 0.556 & $-0.963$ & Collapse \\
Babylonian base-60   & Math     & 0.719 & 0.707 & 0.011 & 0.281 & $-0.346$ & Watch    \\
Schumann resonance   & Physical & 0.651 & 0.543 & 0.108 & 0.349 & $-0.610$ & Collapse \\
Photosynthetic QC    & Physical & 0.494 & 0.379 & 0.115 & 0.506 & $-0.969$ & Collapse \\
\end{tabular}
\end{ruledtabular}
\end{table*}

% ════════════════════════════════════════════════════════════
\section{Conclusion}
\label{sec:conclusion}
% ════════════════════════════════════════════════════════════

Jackson's $\xiJ = 7.2$ is a harmonic ratio in base-60 arithmetic---not a
universal consciousness constant.  When his claims are corrected and
grounded in the GCD kernel, seven theorems emerge, all of which are
computationally verified:

\begin{itemize}
  \item The constant is not privileged ($\rho = +0.25$, not significant).
  \item Recursion does not predict return ($\rho = -0.10$, negative).
  \item Consciousness systems are structurally unstable (0/20 Stable).
  \item One dead channel kills coherence (geometric slaughter).
  \item 432\,Hz equals 440\,Hz in the kernel.
  \item Mathematical identity beats harmonic ratio.
  \item The heterogeneity gap inverts Jackson's hierarchy: biology beats
        harmonics.
\end{itemize}

The salvageable elements---recursive self-reference as a channel, return
through collapse, the heterogeneity gap as ``imperfection,'' and
cross-domain comparison---are precisely the features that map naturally
to GCD structure.  They require only a frozen contract, measured trace
vectors, and seam closure to become welds.

\emph{Quid supersit post collapsum?}  Not the constant.  The questions.

\bibliography{Bibliography}

\end{document}
