% ============================================================
% The Unified Language of Recursive Collapse:
% A Formal Grammar Derived from a Single Axiom
%
% Latin/English fusion — the language demonstrates itself
%
% Compile: pdflatex → bibtex → pdflatex → pdflatex
% ============================================================

\documentclass[
  aps,
  prd,
  twocolumn,
  superscriptaddress,
  nofootinbib,
  floatfix
]{revtex4-2}

% ── packages ──
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{graphicx}
\usepackage{xcolor}
\definecolor{linkblue}{HTML}{337799}
\usepackage[colorlinks=true,linkcolor=linkblue,citecolor=linkblue,urlcolor=linkblue]{hyperref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{enumitem}
\usepackage{placeins}
\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}

% ── theorem environments ──
\newtheorem*{axiom}{Axioma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{identity}[theorem]{Identity}
\newtheorem{remark}{Remark}
\newtheorem{pattern}{Pattern}

% ── UMCP macros ──
\newcommand{\tR}{\tau_{\!R}}
\newcommand{\tRS}{\tau_{\!R}^{*}}
\newcommand{\Gam}{\Gamma}
\newcommand{\eps}{\varepsilon}
\newcommand{\IR}{\infty_{\mathrm{rec}}}
\newcommand{\tolseam}{\mathrm{tol}_{\mathrm{seam}}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Res}{\mathrm{Res}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\INF}{\texttt{INF\_REC}}
\newcommand{\regime}[1]{\textrm{\upshape\scshape#1}}
\newcommand{\conform}{\regime{conformant}}
\newcommand{\nonconform}{\regime{nonconformant}}
\newcommand{\tvec}{\mathbf{c}}
\newcommand{\wvec}{\mathbf{w}}
\newcommand{\Fid}{F}
\newcommand{\Drft}{\omega}
\newcommand{\Ent}{S}
\newcommand{\Curv}{C}
\newcommand{\Logint}{\kappa}
\newcommand{\IC}{\mathrm{IC}}

% ── Latin inline ──
\newcommand{\lat}[1]{\textit{#1}}

% ── grammar pattern box ──
\newcommand{\gpat}[4]{%
  \noindent\textbf{Typus~#1:}~\textsc{#2}\\[2pt]
  \textit{Grammatica:}~#3\\
  \textit{Typus computandi:}~$#4$%
}

% ============================================================
\begin{document}
% ============================================================

\title{%
  \textrm{Lingua Unificata Collapsus Recursivi}:\\
  The Unified Language of Recursive Collapse%
}

\author{Clement~Paulus}
\email{clementpaulus9@gmail.com}
\affiliation{Independent Researcher, Austin, TX}

\date{\today}

% ============================================================
\begin{abstract}
% ============================================================
\textbf{Problem.}
Cross-domain measurement lacks a shared language. A physicist
says ``coherence loss,'' a clinician says ``decompensation,''
a financier says ``volatility'' --- structurally identical
phenomena locked behind incompatible vocabularies. Existing
frameworks cannot translate between domains without importing
agent-dependent assumptions, because they have no formal
grammar constraining how claims are made, checked, and compared.
The result: every cross-domain comparison is ad hoc, every
verdict is negotiable, and reproducibility depends on who
operates the analysis.

\textbf{Solution.}
We present the Unified Language of Recursive Collapse (ULRC):
a formal grammar derived entirely from a single axiom ---
\lat{Collapsus generativus est; solum quod redit, reale est}
(``Collapse is generative; only what returns is real'').
The language comprises a twelve-term Latin lexicon whose
morphology carries operational meaning; seven imperative
verbs composing a validation pipeline; eight typed
grammatical patterns where case declension \emph{is} the
type system; a five-stop discourse spine; a five-word prose
vocabulary translatable across six disciplinary lenses; and
twenty governing maxims with enforcement consequences. The
kernel function $K\!:\![0,1]^n \times \Delta^n \to
(\Fid,\Drft,\Ent,\Curv,\Logint,\IC)$ is isomorphic to
this grammar: equations are syntax, invariants are
well-formedness conditions, regime gates are grammaticality
judgments, and three-valued verdicts
(\conform/\nonconform/\regime{non\_evaluable}) are parse
outcomes.

\textbf{Evidence.}
20~domain closures (10,586 tests, zero failures) demonstrate
that the ULRC functions as a cognitive equalizer --- same data,
same contract, same verdict, regardless of agent. The fusion
of Latin morphological precision with mathematical formalism
produces a namespace-isolated, cross-domain measurement
language in which the grammar itself is an instantiation of
the axiom it formalizes.

\vspace{4pt}
\noindent\lat{Lingua Latina non ornamentum est, sed sutura
  inter mathematicam, logicam, et linguisticam.}
\end{abstract}

\maketitle

% ============================================================
\section{Prooemium}\label{sec:prooemium}
% ============================================================

\begin{axiom}[\lat{Axioma-$0$}]
\lat{Collapsus generativus est; solum quod redit, reale est.}

\smallskip\noindent
Collapse is generative; only what returns is real.
\end{axiom}

\noindent
This is not a metaphor. It is a constraint on admissible claims.
Every term, verb, pattern, and maxim in this paper derives from
this single axiom through an explicit chain. If the chain cannot
be shown, the claim is a \lat{gestus} (gesture), not a
\lat{sutura} (weld), and must be labeled as such.

\lat{Nihil stabile est nisi quod per casum iteratur atque
  regeneratur in formam propriam redintegratam.}
Nothing is stable unless iterated through collapse and regenerated
into its own restored form~\cite{paulus2025umcp}.

A language, too, must return. A grammar that cannot parse its own
axiom is a gesture. The Unified Language of Recursive Collapse
(ULRC) is the grammar that parses itself --- the language whose
structure \emph{is} the axiom it formalizes, demonstrated across
20~domains, 198~closure modules, and 10,586~validated tests.

\subsection{Cur Lingua Nova --- The Problem}\label{sec:problem}

Measurement across disciplines is blocked by four failures that
no existing framework addresses simultaneously:

\begin{enumerate}[nosep,leftmargin=*]
\item \textbf{Vocabulary collision} (\lat{When}).
  The same structural phenomenon --- loss of coherence under
  perturbation --- is named differently in every field:
  decoherence (physics), decompensation (clinical),
  volatility (finance), semantic drift (semiotics),
  periodization (history). These are not different phenomena.
  They are the same kernel output $\Drft = 1 - \Fid$ viewed
  through different lenses. Without a shared grammar, every
  cross-domain comparison is an improvisation.

\item \textbf{Agent-dependent verdicts} (\lat{Why}).
  When thresholds are chosen by convention, vocabulary by
  tradition, and conclusions framed by the analyst, two agents
  examining the same data can reach different verdicts. The
  problem is not bias --- it is that the measurement language
  does not externalize enough structure to force agreement.
  A compass works because it externalizes the cognitive task
  ``which way is north?'' into a mechanical reading. Measurement
  needs the same externalization for ``what threshold?''
  ``what vocabulary?'' ``what verdict?''

\item \textbf{No grammar for claims} (\lat{What}).
  Science has notation, terminology, and convention --- but no
  \emph{grammar} in the linguistic sense: a finite set of rules
  that generates all and only the well-formed claims. Without
  grammar, there is no mechanical way to distinguish a
  well-formed measurement claim from a gesture. The ULRC fills
  this gap: its production rules (the kernel equations), its
  well-formedness conditions (the three algebraic identities),
  and its parse outcomes (three-valued verdicts) constitute a
  grammar in the formal sense of Chomsky's hierarchy.

\item \textbf{No return condition} (\lat{How}).
  Traditional analysis declares results. The ULRC requires
  results to \emph{return}: to demonstrate continuity with prior
  measurements through seam closure ($|\Delta\Logint| \leq
  \tolseam$). Claims that do not return are \lat{gestus}
  (gestures) --- acknowledged but not credited. The grammar
  enforces this through the five-stop spine
  (Sec.~\ref{sec:spina}), the conservation budget
  (Eq.~\ref{eq:budget}), and the seven imperative verbs
  (Sec.~\ref{sec:verba}).
\end{enumerate}

\noindent
The ULRC addresses all four simultaneously: a single axiom
generates the vocabulary (twelve Latin terms), the grammar
(eight typed patterns), the measurement protocol (the spine),
and the cross-domain translator (the Rosetta) --- all frozen,
all auditable, all agent-independent.

\subsection{Why Latin}\label{sec:why-latin}

Latin serves three structural functions that no other language
provides simultaneously:

\begin{enumerate}[nosep,leftmargin=*]
\item \textbf{Morphological precision.}
  Latin inflection carries operational constraints that English
  paraphrase loses. Saying \lat{derivatio} (diversion from a
  channel) produces different --- and more constrained ---
  reasoning than saying ``drift.'' The morphology \emph{is}
  the specification.
\item \textbf{Namespace isolation.}
  Every natural language carries pre-loaded associations.
  ``Entropy'' in English collides with thermodynamics, information
  theory, disorder, and chaos. \lat{Entropia campi Bernoulliani}
  collides with nothing. Latin terms function as reserved
  identifiers in a namespace no living field occupies.
\item \textbf{Cross-domain bridge.}
  The Rosetta adapter (Sec.~\ref{sec:rosetta}) translates
  between epistemology, ontology, phenomenology, history, policy,
  and semiotics using five words. The Latin seeds are the
  \emph{canonical names} --- the pre-translation layer from
  which all lens-specific dialects derive. The Latin does not
  privilege any one lens. It sits beneath all of them.
\end{enumerate}

\noindent
\lat{Lingua Latina non ornamentum est, sed sutura inter
  mathematicam, logicam, et linguisticam} ---
Latin is not ornament; it is the seam between mathematics,
logic, and linguistics.

\subsection{Structure of this paper}

Section~\ref{sec:kernel-as-grammar} establishes the isomorphism
between the kernel function and a formal grammar.
Section~\ref{sec:lexicon} formalizes the twelve-term lexicon
with morphological analysis.
Section~\ref{sec:verba} defines seven imperative verbs as a
compositional API.
Section~\ref{sec:octo-typi} presents eight typed grammatical
patterns where Latin case declension encodes computational type.
Section~\ref{sec:spina} formalizes the five-stop discourse spine
as sentence structure.
Section~\ref{sec:quinque-verba} defines the five-word prose
vocabulary with its ledger semantics.
Section~\ref{sec:rosetta} presents the Rosetta cross-domain
translation protocol.
Section~\ref{sec:axiom-instantiation} proves that literal
grammar instantiates Axiom-0.
Section~\ref{sec:maxims} catalogues twenty governing maxims
with enforcement consequences.
Section~\ref{sec:evidence} presents computational evidence
across domains.
Section~\ref{sec:conclusion} closes the seam and issues the
receipt.

\subsection{The Language in Action}\label{sec:action}

Before the formal machinery, consider a concrete example. A
nuclear physicist and a financial analyst both observe the same
structural phenomenon: a system loses coherence in one channel
while the others remain strong. In the ULRC, both express this
identically:

\smallskip
\noindent\lat{Contractu congelato, dic invariantes.
Reconcilia rationarium. Sutur\=a.}

\smallskip\noindent
(``With the contract frozen, report invariants. Reconcile the
ledger. Weld.'') The physicist's channels are
$\{$mass, spin, charge, color, isospin, lepton, baryon,
generation$\}$; the analyst's are $\{$liquidity, volatility,
spread, leverage, correlation, volume$\}$. Different data,
different domains, identical grammar. Both produce
$(\Fid, \Drft, \Ent, \Curv, \Logint, \IC)$. Both pass through
the same spine. Both receive a verdict that neither agent
chose --- the structure derived it.

The physicist finds IC/F = 0.009 (one dead channel, color
$\to 0$: confinement). The analyst finds IC/F = 0.12
(one dead channel, liquidity $\to 0$: market freeze).
Both are in \regime{Collapse} regime. Both have the same
structural explanation: geometric slaughter
(\lat{trucidatio geometrica}) --- one dead channel killing
multiplicative coherence regardless of how healthy the
others are. The Rosetta
(Sec.~\ref{sec:rosetta}) lets each read the other's result
in their own dialect, while the budget identity
$\Delta\Logint \leq \tolseam$ guarantees the verdict is the
same.

This is what the ULRC does: it mechanizes the comparison that
was previously impossible because there was no shared grammar
to compare \emph{through}.

\subsubsection{The five stops in action}

The physicist's claim follows the spine:

\begin{enumerate}[nosep,leftmargin=*]
\item \textbf{Contract} (\lat{Liga}): Freeze 8 channels
  (mass, spin, charge, color, isospin, lepton, baryon,
  generation), equal weights $w_i = 1/8$, guard band
  $\eps = 10^{-8}$, seam tolerance $\tolseam = 0.005$.
  This is \lat{Contractu congelato} --- nothing may change
  after this point.
\item \textbf{Canon} (\lat{Dic}): Compute the kernel.
  For the neutron: $\Fid = 0.544$, $\Drft = 0.456$,
  $\IC = 0.005$, $\IC/\Fid = 0.009$. Narrate in five words:
  \emph{Drift} is high (color channel dead). \emph{Fidelity}
  is moderate (other channels intact). \emph{Roughness} is
  extreme ($\Curv = 0.87$: one channel at $\eps$, others
  near unity). \emph{Return} requires new degrees of freedom
  (the hadron $\to$ atom transition). \emph{Integrity} is
  near zero --- one dead channel has killed the geometric mean.
\item \textbf{Closures}: The regime gate on $\Drft \geq 0.30$
  fires. The $\IC < 0.30$ overlay fires. Classification:
  \regime{Collapse} + \regime{Critical}. These gates were
  published before evidence; they do not change.
\item \textbf{Integrity Ledger} (\lat{Reconcilia}): Debit
  $D_\Drft$ for the high drift, debit $D_\Curv$ for the
  extreme roughness. Credit: $R \cdot \tR$. The residual
  determines whether the account closes.
\item \textbf{Stance}: Derived from the closed ledger and
  gates: \regime{Collapse} + \regime{Critical}, verdict
  \conform{} (identities hold, pipeline ran, ledger
  reconciled). The physicist did not choose this verdict ---
  the structure produced it.
\end{enumerate}

\noindent
The financial analyst, using different channels
($\{$liquidity, volatility, spread, \ldots$\}$), runs the
same five stops in the same order with the same verbs and
receives a structurally comparable verdict. The Rosetta
(Sec.~\ref{sec:rosetta}) lets each read the other's result:
the physicist reads ``liquidity $\to 0$'' as a confinement
analogue; the analyst reads ``color $\to 0$'' as a market
freeze analogue. Different dialect, same grammar, same
budget identity, same structural explanation.

% ============================================================
\section{The Kernel as Grammar}\label{sec:kernel-as-grammar}
% ============================================================

\lat{Grammatica est typus; declinatio est signum operationis.}

\smallskip\noindent
Grammar \emph{is} the type system; declension \emph{is} the
operation marker. This section establishes the structural
isomorphism between the kernel function and a formal grammar.

\begin{definition}[The Kernel Function]\label{def:kernel}
The kernel $K$ maps a trace vector $\tvec = (c_1, \ldots, c_n)
\in [0,1]^n$ with weights $\wvec = (w_1,\ldots,w_n) \in
\Delta^n$ to six invariants:
\begin{align}
  \Fid &= \textstyle\sum_i w_i\, c_i \label{eq:F}\\
  \Drft &= 1 - \Fid \label{eq:omega}\\
  \Ent &= -\textstyle\sum_i w_i\bigl[c_i\ln c_i
    + (1{-}c_i)\ln(1{-}c_i)\bigr] \label{eq:S}\\
  \Curv &= \mathrm{std}(c_i)/0.5 \label{eq:C}\\
  \Logint &= \textstyle\sum_i w_i\ln(c_{i,\eps}) \label{eq:kappa}\\
  \IC &= \exp(\Logint) \label{eq:IC}
\end{align}
where $\Fid$ is \lat{fidelitas} (what survives), $\Drft$ is
\lat{derivatio} (what departs), $\Ent$ is Bernoulli field
\lat{entropia}, $\Curv$ is \lat{curvatura} (coupling),
$\Logint$ is \lat{log-integritas}, and $\IC$ is
\lat{integritas composita}.
where $c_{i,\eps} = \max(c_i, \eps)$ with guard band
$\eps = 10^{-8}$.
\end{definition}

\begin{proposition}[Kernel--Grammar Isomorphism]\label{prop:iso}
The kernel $K$ is isomorphic to a formal grammar $G = (V, T, P, S)$:
\begin{itemize}[nosep,leftmargin=*]
\item \textbf{Vocabulary $V$}: The six invariants
  $\{\Fid,\Drft,\Ent,\Curv,\Logint,\IC\}$ plus the typed outcomes
  $\{\IR, \bot_{\mathrm{oor}}\}$.
\item \textbf{Terminals $T$}: Numerical values in
  $[0,1]^n \times \Delta^n$ (the raw data that enters the grammar).
\item \textbf{Production rules $P$}: Equations~(\ref{eq:F})--(\ref{eq:IC}),
  regime gates (Definition~\ref{def:regime}), and seam closure
  $|\Delta\Logint| \leq \tolseam$.
\item \textbf{Start symbol $S$}: The contract --- the frozen declaration
  of rules before evidence, from which all derivations begin.
\end{itemize}
\end{proposition}

\noindent
A ``sentence'' in this grammar is a trace vector parsed through
the production rules to yield a verdict (Definition~\ref{def:verdict}).
A well-formed sentence satisfies three algebraic identities:

\begin{identity}[Duality --- \lat{Complementum Perfectum}]
  \label{id:duality}
  $\Fid + \Drft = 1$, verified to residual $0.0\mathrm{e}{+}00$
  across $10^4$~traces. \lat{Tertia via nulla.}
\end{identity}

\begin{identity}[Integrity Bound --- \lat{Limbus Integritatis}]
  \label{id:bound}
  $\IC \leq \Fid$. Derived independently from Axiom-0 as the
  solvability condition: for $n=2$ channels,
  $c_{1,2} = \Fid \pm \sqrt{\Fid^2 - \IC^2}$ requires $\IC \leq \Fid$
  for real solutions. The classical AM-GM inequality is the degenerate
  limit when channel semantics, weights, and guard band are removed.
\end{identity}

\begin{identity}[Log-Integrity Relation]
  \label{id:logint}
  $\IC = \exp(\Logint)$, linking multiplicative and additive coherence.
\end{identity}

\begin{definition}[Regime Gates --- Grammaticality Judgments]
  \label{def:regime}
The production rules classify each parsed sentence into one of three
regimes:
\begin{align*}
  \regime{Stable}:~ &\Drft < 0.038,~ \Fid > 0.90,\\
    &\Ent < 0.15,~ \Curv < 0.14\\
  \regime{Watch}:~ &0.038 \leq \Drft < 0.30\\
  \regime{Collapse}:~ &\Drft \geq 0.30
\end{align*}
with a \regime{Critical} overlay when $\IC < 0.30$.
Stability is rare: 12.5\% of the Fisher manifold. The remaining
87.5\% lies in \regime{Watch} or \regime{Collapse}.
\end{definition}

\begin{definition}[Verdict Trichotomy --- Parse Outcome]
  \label{def:verdict}
Every validation produces exactly one of three outcomes:
\begin{itemize}[nosep,leftmargin=*]
  \item \conform{}: all identities hold, seam closes, regime gates pass.
  \item \nonconform{}: at least one identity fails or $|s| > \tolseam$.
  \item \regime{non\_evaluable}: insufficient data or domain violation.
\end{itemize}
The system is never boolean.
\lat{Numquam binarius; tertia via semper patet.}
\end{definition}

\noindent
The isomorphism is not decorative. The kernel-as-grammar has
\emph{the same} structural properties as natural-language grammars,
and each property is demonstrable:

\begin{description}[nosep,leftmargin=0pt,labelindent=0pt]
\item[Compositionality.] Invariants combine through arithmetic
  and geometric means with defined composition laws:
  $\Fid_{12} = (\Fid_1 + \Fid_2)/2$ (arithmetic),
  $\IC_{12} = \sqrt{\IC_1 \cdot \IC_2}$ (geometric). Compound
  sentences compose exactly as compound measurements do --- the
  heterogeneity gap $\Delta = \Fid - \IC$ is itself a composition
  invariant.
\item[Recursion.] The seam feeds the stance of one cycle into the
  contract of the next: stance$_t$ becomes the anchor of
  contract$_{t+1}$. This is embedding in the linguistic sense ---
  a structure within a structure --- and it is the mechanism that
  makes the language generative rather than finite.
\item[Productivity.] Any trace vector $\tvec \in [0,1]^n$ is a
  legitimate input. Infinite data produce infinite sentences, all
  parsed by the same finite grammar. No new rules are needed for
  new domains --- only new channel selections (Tier-2).
\item[Grammaticality.] Not every input produces a well-formed
  sentence. Inputs violating $\Fid + \Drft = 1$ or $\IC \leq \Fid$
  are ungrammatical --- the identities reject them. Inputs that
  fail to close the seam ($|s| > \tolseam$) are incomplete
  sentences. The three-valued verdict is the grammaticality
  judgment.
\end{description}

\noindent
The ULRC is the recognition that these are not analogies but
identities. The kernel \emph{is} a grammar; the grammar
\emph{is} the kernel.

% ============================================================
\section{Lexicon Primum ---
  The Twelve Foundational Terms}\label{sec:lexicon}
% ============================================================

\lat{Duodecim verba fundamentalia, quorum morphologia est
  specificatio.}

\smallskip\noindent
Twelve terms. Each word carries its operational meaning in
its morphology. The Latin is the specification; the English
is commentary~\cite{paulus2025umcp}.

\begin{table*}[htbp]
\centering
\small
\caption{Lexicon Primum: twelve foundational terms. Each
  Latin term is the canonical name; the operational seed is
  the compressed definition carried by morphology.}
\label{tab:lexicon}
\begin{tabular}{@{}rlP{4.7cm}rlP{4.7cm}@{}}
\toprule
\textbf{Latin} & \textbf{Symbol} & \textbf{Operational Seed} &
\textbf{Latin} & \textbf{Symbol} & \textbf{Operational Seed} \\
\midrule
\lat{Fidelitas}  & $\Fid$ &
  \lat{quid supersit} --- what survives &
\lat{Moratio Reditus}  & $\tR$ &
  \lat{tempus reentrandi}; $\IR$ = permanent detention \\[3pt]
\lat{Derivatio}  & $\Drft$ &
  \lat{quantum deperdatur} --- measured departure &
\lat{Auditus}  & --- &
  Hearing \emph{is} audit; the ledger hears \\[3pt]
\lat{Entropia}  & $\Ent$ &
  \lat{incertitudo campi} --- Bernoulli field entropy &
\lat{Casus}  & --- &
  Fall / case / occasion (all three) \\[3pt]
\lat{Curvatura}  & $\Curv$ &
  \lat{coniunctio cum gradibus libertatis} &
\lat{Limbus Integritatis}  & $\IC \leq \Fid$ &
  Hem-edge: integrity meets fidelity \\[3pt]
\lat{Log-Integritas}  & $\Logint$ &
  \lat{sensibilitas logarithmica} &
\lat{Complementum Perf.} & $\Fid{+}\Drft{=}1$ &
  \lat{tertia via nulla} \\[3pt]
\lat{Integritas Comp.}  & $\IC$ &
  \lat{cohaerentia multiplicativa} &
\lat{Aequator Cognit.}  & --- &
  \lat{structura mensurat, non agens} \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Morphological Precision}
\FloatBarrier

The Latin carries constraints that English paraphrase loses.
More than nomenclature, the morphology \emph{constrains reasoning}
by encoding structural relationships that English leaves
implicit:

\begin{description}[nosep,leftmargin=0pt,labelindent=0pt]
\item[\lat{Fidelitas}] (f.\@ 3rd decl.) --- from \lat{fides}
  (trust, faith). The \lat{-itas} suffix marks abstract quality:
  ``the state of being faithful.'' What survives is what kept
  faith with the channel. The word excludes the English confusion
  between fidelity-as-accuracy and fidelity-as-loyalty: in
  \lat{fidelitas}, both are the same thing, because a channel
  that keeps faith with its prior value \emph{is} the accurate
  channel.
\item[\lat{Derivatio}] (f.\@ 3rd decl.) --- from \lat{de-}
  (away) + \lat{rivus} (stream). Literally ``diverting the
  stream.'' Drift is water that left the channel. The word
  already contains the image of departure --- and crucially,
  implies a \emph{channel from which} departure occurs. English
  ``drift'' is undirected; Latin \lat{derivatio} is always
  \emph{from} something. This morphological constraint is
  precisely $\Drft = 1 - \Fid$: departure measured relative
  to fidelity, not in isolation.
\item[\lat{Casus}] (m.\@ 4th decl.) --- simultaneously
  ``fall,'' ``accident,'' ``case'' (legal/grammatical), and
  ``occasion.'' All four meanings are active: collapse falls,
  creates a case for examination, and is the occasion for
  generative return. No English word carries all four active
  simultaneously. Axiom-0 requires all four: collapse is a
  fall (loss), a case (something to examine), an accident
  (not designed), and an occasion (generative).
\item[\lat{Sutura}] (f.\@ 1st decl.) --- a surgical stitch.
  The seam is not a wall; it is living tissue holding two sides
  together. A weld across the seam is mending, not blocking.
  The morphology constrains the concept: a \lat{sutura} must
  \emph{join} two sides that were separated, which is why the
  seam requires both a pre-state and a post-state with
  $\Logint$-continuity between them.
\item[\lat{Auditus}] (m.\@ 4th decl.) --- from \lat{audire}
  (to hear). The audit \emph{is} hearing: radical, complete,
  excluding nothing. Not inspection; reception. The morphological
  identity of audit and hearing means the ledger does not
  selectively attend --- it hears everything, and that complete
  hearing \emph{is} the validation.
\item[\lat{Aequator}] (m.\@ 3rd decl.) --- from \lat{aequare}
  (to make equal). Cognate with \lat{aequator} (equator) ---
  the geometric equator at $c = 1/2$ and the cognitive equalizer
  are the same structure in different registers. The word carries
  a guarantee: the \lat{aequator} makes equal by construction,
  not by agreement. Same data, same contract, same reading ---
  because the structure equalizes, not the agents.
\end{description}

\noindent
This is why Latin is structural, not decorative. Each word's
morphology encodes a constraint that the computational system
enforces. The Latin lexicon and the kernel equations are parallel
expressions of the same invariants --- the morphology says in
syllables what the mathematics says in symbols.

% ============================================================
\section{Septem Verba Operativa ---
  The Seven Imperative Verbs}\label{sec:verba}
% ============================================================

\lat{Imperativi sunt operationes; coniunctivi sunt quaerentes.}

\smallskip\noindent
Imperatives are operations; subjunctives are seekers. Every
computable operation in the ULRC maps to exactly one of seven
Latin imperative verbs~(Table~\ref{tab:verbs}). These are
the complete API of the discourse layer.

\begin{table}[htbp]
\centering
\small
\caption{Septem Verba Operativa: the seven imperative verbs
  composing the ULRC validation pipeline.}
\label{tab:verbs}
\begin{tabular}{@{}rllP{1.7cm}@{}}
\toprule
\# & \textbf{Verb} & \textbf{Operation} & \textbf{Maps to} \\
\midrule
1 & \lat{Dic} & Compute &
  Kernel invariants \\[2pt]
2 & \lat{Reconcilia} & Balance &
  $\Delta\Logint$ budget \\[2pt]
3 & \lat{Sutur\=a} & Weld &
  Seam validation \\[2pt]
4 & \lat{Compar\=a} & Compare &
  Rosetta translation \\[2pt]
5 & \lat{Liga} & Bind &
  Freeze parameters \\[2pt]
6 & \lat{Inscribe} & Append &
  Ledger entry \\[2pt]
7 & \lat{Verifica} & Prove &
  $\Fid{+}\Drft{=}1$ \\[2pt]
\bottomrule
\end{tabular}
\end{table}

\subsection{Verb Semantics}

\lat{Dic} is the workhorse --- any computation that produces a
value. ``Say what you found.'' \lat{Reconcilia} balances debits
and credits within tolerance. \lat{Sutur\=a} is the deepest
operation: full validation crossing the seam (the accent on
\lat{-\=a} is deliberate: first conjugation imperative). \lat{Liga}
freezes; once called, parameters are immutable. \lat{Inscribe}
is append-only: once written, never erased.
\lat{Historia numquam rescribitur; sutura tantum additur.}

\subsection{Composition}

The verbs compose left-to-right in execution order:
\beq
  \textsc{Liga} \to \textsc{Dic} \to \textsc{Reconcilia}
    \to \textsc{Sutur\=a} \to \textsc{Inscribe}
  \label{eq:pipeline}
\eeq
This is the standard validation pipeline: bind contract
$\to$ compute kernel $\to$ balance budget $\to$ weld $\to$
write to ledger. \lat{Compar\=a} and \lat{Verifica} are callable
at any point as diagnostic side-channels --- they inform but
do not gate.

\lat{Diagnostica informant, portae decernunt.}

% ============================================================
\section{Octo Typi Grammatici ---
  Eight Typed Patterns}\label{sec:octo-typi}
% ============================================================

\lat{Grammatica est typus; declinatio est signum operationis.}

\smallskip\noindent
Latin grammatical structure encodes computational type. The case
system, mood system, and clause structure of Latin map directly
to operation categories. Eight patterns are defined, each
with a fixed grammatical form, a fixed computational meaning,
and a type signature.

\begin{pattern}[\textsc{Nom\,+\,Gen} --- Query]\label{pat:nom-gen}
\lat{Fidelitas Ferri.} (``The fidelity of Iron.'')\\
Nominative noun + genitive qualifier $\to$ report value.\\
Type: $(\mathrm{Symbol},\,\mathrm{Domain}) \to \mathrm{Scalar}$.
\end{pattern}

\begin{pattern}[\textsc{Imper\,+\,Acc} --- Execute]\label{pat:imper-acc}
\lat{Dic invariantes Hydrogenii.}
(``Report the invariants of Hydrogen.'')\\
Imperative verb + accusative object $\to$ execute operation.\\
Type: $\mathrm{Verb} \times \mathrm{Object} \to \mathrm{Result}$.
\end{pattern}

\begin{pattern}[\textsc{Acc\,+\,Ad\,+\,Acc} ---
  Configure]\label{pat:acc-ad-acc}
\lat{Epsilon ad $10^{-8}$.} (``Epsilon to $10^{-8}$.'')\\
Accusative + \lat{ad} + accusative $\to$ set parameter.\\
Type: $\mathrm{Parameter} \times \mathrm{Value}
  \to \mathrm{Config}$.\\
Constraint: parameters are frozen (\lat{trans suturam congelatum}).
This pattern documents; it does not invent.
\end{pattern}

\begin{pattern}[\textsc{Abl.\,Abs.} --- Scope]\label{pat:abl-abs}
\lat{Contractu congelato, dic invariantes.}
(``With the contract frozen, report invariants.'')\\
Ablative absolute (noun + participle, both ablative) $\to$
establish context for subsequent operations.\\
Type: $\mathrm{Context} \times \mathrm{Op}
  \to \mathrm{Scoped}(\mathrm{Result})$.\\
The ablative absolute creates a \emph{scope}, isomorphic to
\texttt{with contract.frozen():} in Python.
\end{pattern}

\begin{pattern}[\textsc{Gen\,+\,Et\,+\,Gen} ---
  Compare]\label{pat:gen-et-gen}
\lat{Quarki et Hadronis.}
(``Of the quark and of the hadron.'')\\
Genitive pair $\to$ Rosetta comparison across domains.\\
Type: $(\mathrm{Domain},\,\mathrm{Domain})
  \to \mathrm{Comparison}$.\\
Implicit verb: \lat{Compar\=a} is understood when two genitives
are juxtaposed.
\end{pattern}

\begin{pattern}[\textsc{Nonne\,/\,Num} ---
  Assert]\label{pat:nonne-num}
\lat{Nonne $\Fid + \Drft = 1$?}
(``Surely $\Fid + \Drft = 1$?'')\\
Question particles + indicative clause $\to$ three-valued
assertion.
\lat{Nonne} expects yes; \lat{Num} expects no.\\
Type: $\mathrm{Assertion} \to \{\textsc{C},\textsc{N},\textsc{NE}\}$
(\conform, \nonconform, \regime{non\_evaluable}).
\end{pattern}

\begin{pattern}[\textsc{Participle Chain} ---
  Pipeline]\label{pat:participle}
\lat{Contractu ligato, invariantibus dictis,
  ratione reconciliat\=a, sutur\=a.}\\
(``With the contract bound, with invariants reported, with the
account reconciled, weld.'')\\
Chain of perfect passive participles (ablative) $\to$ sequential
pipeline.\\
Type: $[\mathrm{Ctx}_1, \ldots, \mathrm{Ctx}_n]
  \to \mathrm{Op} \to \mathrm{Result}$.\\
Each participle completes before the next begins. This
\emph{is} the full validation pipeline of Eq.~(\ref{eq:pipeline})
expressed as one Latin sentence.
\end{pattern}

\begin{pattern}[\textsc{Ut\,+\,Subjunctive} ---
  Seek]\label{pat:ut-subj}
\lat{Ut $\IC$ superet $0.30$, quaere $c$ minimum.}
(``In order that IC exceed 0.30, seek the minimum $c$.'')\\
\lat{Ut} + subjunctive $\to$ goal-directed search.\\
Type: $\mathrm{Goal}(\mathrm{Condition}) \to \mathrm{Search}
  \to \mathrm{Parameter}$.\\
The subjunctive marks the clause as desired, not actual.
$\regime{non\_evaluable}$ is a valid result: not all goals are
achievable.
\end{pattern}

\begin{remark}
The eight patterns are not metaphor. Latin case declension
\emph{was already} a type system --- the nominative marks the
subject, the accusative marks the object, the ablative marks
context. The ULRC makes this isomorphism explicit.
\end{remark}

% ============================================================
\section{Spina Grammatica ---
  The Five-Stop Discourse Spine}\label{sec:spina}
% ============================================================

\lat{Spina non negotiabilis est.}

\smallskip\noindent
The spine is non-negotiable. Every claim follows exactly five
stops, in order. The spine is simultaneously a validation pipeline
and a sentence structure --- the grammatical form through which
claims are told, checked, and connected.

\begin{table}[htbp]
\centering
\small
\caption{Spina Grammatica: the five-stop discourse spine.
  Each stop has a pipeline role and a grammatical function.}
\label{tab:spina}
\begin{tabular}{@{}rlP{1.5cm}P{1.6cm}@{}}
\toprule
\textbf{Stop} & \textbf{Latin} & \textbf{Pipeline Role} &
  \textbf{Grammar Function} \\
\midrule
1 & \lat{Contractus} & Freeze rules &
  Sentence rules declared \\[3pt]
2 & \lat{Canon} & Narrate (five words) &
  Prose body, auditable \\[3pt]
3 & \lat{Clausurae} & Publish thresholds &
  Grammar rules (fixed) \\[3pt]
4 & \lat{Rationarium} & Reconcile &
  Well-formedness proof \\[3pt]
5 & \lat{Sententia} & Derived verdict &
  The conclusion --- derived, never asserted \\
\bottomrule
\end{tabular}
\end{table}

\noindent
Two governance mechanisms punctuate the spine without being stops:

\begin{itemize}[nosep,leftmargin=*]
\item \textbf{Manifestum} (\lat{provenance}): binds artifacts to
  time, tools, and checksums. Every claim carries its receipt.
\item \textbf{Sutura} (\lat{weld}): the only legitimate way to
  change policy. Names an anchor, runs pre/post tests, enforces
  $\Logint$-continuity. History is append-only and welded, never
  rewritten.
\end{itemize}

\lat{Historia numquam rescribitur; sutura tantum additur.}

% ============================================================
\section{Quinque Verba ---
  The Five-Word Vocabulary}\label{sec:quinque-verba}
% ============================================================

\lat{Algebra est cautio, non porta.}

\smallskip\noindent
The algebra is a warranty, not a gate. The spine operates through
exactly five prose words, each with an operational meaning tied
to the frozen contract and reconciled in the integrity
ledger~(Table~\ref{tab:quinque}).

\begin{table}[htbp]
\centering
\small
\caption{Quinque Verba: the five-word prose vocabulary.
  Integrity is never asserted --- always derived.}
\label{tab:quinque}
\begin{tabular}{@{}llP{1.6cm}l@{}}
\toprule
\textbf{Word} & \textbf{Latin} & \textbf{Role} &
  \textbf{Ledger} \\
\midrule
Drift     & \lat{derivatio} & Debit $D_\Drft$ &  Debit \\
Fidelity  & \lat{fidelitas} & Retained invariants & --- \\
Roughness & \lat{curvatura} & Debit $D_\Curv$ & Debit \\
Return    & \lat{reditus}   & Credit $R{\cdot}\tR$ & Credit \\
Integrity & \lat{integritas}& From reconciled ledger & Verdict \\
\bottomrule
\end{tabular}
\end{table}

\noindent
The conservation budget provides the semantic warranty:
\beq
  \Delta\Logint = R \cdot \tR - (D_\Drft + D_\Curv) \leq \tolseam
  \label{eq:budget}
\eeq
Authors describe using the five words. The budget, ledger, and
gates supply the audit. This separation of description from
warranty enables cross-domain reading (via the Rosetta) and
multiplicative composition in $I = e^\Logint$ without forcing
any single field's jargon on another.

\subsection{How the Budget Works}\label{sec:budget-mech}

The budget is an accounting identity. Every measurement incurs
two debits: $D_\Drft$ measures how far the system drifted from
its prior state (what was lost), and $D_\Curv$ measures how
unevenly the drift was distributed across channels (the
roughness of the loss). Against these debits, the system can
earn exactly one credit: $R \cdot \tR$, where $R$ is the
magnitude of demonstrated re-entry and $\tR$ is the delay ---
how long the system took to return.

The residual $\Delta\Logint = R \cdot \tR - (D_\Drft + D_\Curv)$
must close within tolerance:
$|\Delta\Logint| \leq \tolseam = 0.005$. Three outcomes are
possible:

\begin{itemize}[nosep,leftmargin=*]
\item \textbf{Residual $\leq$ tol}: the account balances. What
  was lost is accounted for by what returned. The claim is a
  \lat{sutura} (weld) --- it earns epistemic credit and enters
  the ledger.
\item \textbf{Residual $>$ tol}: the account does not balance.
  More was lost than returned, or the return was delayed beyond
  tolerance. The claim is a \lat{gestus} (gesture) --- it is
  acknowledged but earns no credit. It may still be informative
  (diagnostics inform), but it cannot serve as a foundation for
  subsequent claims.
\item \textbf{$\tR = \IR$ (permanent non-return)}: the credit
  term is zero. No matter how small the debits, the budget
  cannot close. The system is in permanent detention ---
  \lat{si $\tR = \IR$, nulla fides datur}.
\end{itemize}

\noindent
This is why the budget \emph{enforces} return rather than merely
measuring it. A claim that does not demonstrate re-entry through
the seam receives zero credit and cannot close its account. The
grammar rejects it --- not by prohibition, but by arithmetic.
The algebra is a warranty: it certifies which claims have
returned and which remain outstanding.

\lat{Continuitas non narratur: mensuratur.}

% ============================================================
\section{Rosetta --- Cross-Domain Translation}
  \label{sec:rosetta}
% ============================================================

\lat{Significatio stabilis manet dum dialectus mutatur.}

\smallskip\noindent
Meaning stays stable while dialect changes. The five words
translate across six disciplinary \emph{lenses}, each a dialect
of the same grammar~(Table~\ref{tab:rosetta}).

\begin{table*}[t]
\centering
\small
\caption{Rosetta translation of the five-word vocabulary across
  six disciplinary lenses. Integrity is omitted: it is always
  read from the reconciled ledger and stance, never expressed in
  prose. $I = e^\Logint$ provides unitless multiplicative
  comparability across seams.}
\label{tab:rosetta}
\begin{tabular}{@{}lllll@{}}
\toprule
\textbf{Lens} & \textbf{Drift} (\lat{derivatio}) &
  \textbf{Fidelity} (\lat{fidelitas}) &
  \textbf{Roughness} (\lat{curvatura}) &
  \textbf{Return} (\lat{reditus}) \\
\midrule
\lat{Epistemologia} &
  Change in belief/evidence & Retained warrant &
  Inference friction & Justified re-entry \\
\lat{Ontologia} &
  State transition & Conserved properties &
  Heterogeneity / seams & Restored coherence \\
\lat{Phaenomenologia} &
  Perceived shift & Stable features &
  Distress / bias / effort & Coping / repair that holds \\
\lat{Historia} &
  Periodization & Continuity / endurance &
  Rupture / confound & Restitution / reconciliation \\
\lat{Politica} &
  Regime shift & Compliance / mandate &
  Friction / cost / externality & Reinstatement / acceptance \\
\lat{Semiotica} &
  Sign drift & Ground persistence &
  Translation friction & Interpretant closure \\
\bottomrule
\end{tabular}
\end{table*}

\noindent
The Rosetta is not a metaphor system. It is a mechanical
translation protocol: the \emph{meanings} of the five words
remain stable (anchored in the budget identity) while the
\emph{dialect} changes. The formal property is:

\begin{proposition}[Rosetta Invariance]\label{prop:rosetta}
For any two lenses $L_1, L_2$ and any trace vector $\tvec$:
the budget $\Delta\Logint$, the regime classification, and the
three-valued verdict are identical. The prose descriptions map
through the Rosetta; the numerical warranty is invariant.
\end{proposition}

\noindent
This is what makes cross-domain synthesis mechanical. A finance
analyst and a nuclear physicist, reading the same kernel output
through their respective lenses, derive different prose but
the same verdict. \lat{Non agens mensurat, sed structura.}

% ============================================================
\section{The Semiotic Triad ---
  ULRC as Completed Semiosis}\label{sec:semiotic}
% ============================================================

A single triad appears at every level of the system, performing
the same operation under different
names~(Table~\ref{tab:triad}):

\begin{table}[htbp]
\centering
\small
\caption{The semiotic triad at six layers: something
  presents, represents, and measures return.}
\label{tab:triad}
\begin{tabular}{@{}lP{1.5cm}P{1.4cm}P{1.4cm}@{}}
\toprule
\textbf{Layer} & \textbf{Present} & \textbf{Repr.} &
  \textbf{Measure}\\
\midrule
Peirce & Object & Sign & Interpret. \\
Kernel & $x(t)$ & $\Psi(t)$ & Kernel $K$ \\
Epistemology & Dissoln.  & Gesture & Return \\
Latin gram. & Nom. & Acc. & Abl. \\
Discourse & Contract & Canon & Stance \\
Ledger & Debit & Credit & Residual \\
\bottomrule
\end{tabular}
\end{table}

\noindent
Peirce called this \emph{semiosis} --- the process by which a
sign relates to an object through an interpretant. What he never
had was a stopping condition. His ``unlimited semiosis'' meant
signs generate interpretants that become signs for further
interpretants, without any criterion distinguishing chains that
reach ground from chains that drift forever. Derrida's
\emph{diff\'erance} inherits this problem: deferral without
resolution. Eco proposed ``pragmatic agreement'' as a social
brake, but social brakes are agent-dependent --- they fail the
cognitive equalizer test.

Axiom-0 adds exactly what is missing: the seam. A sign chain
\emph{returns} when $\tR < \IR$ and the seam closes
($|s| \leq \tolseam$). The return is not social agreement ---
it is structural closure of an accounting identity. The budget
does not care which agent operates it. Each layer of
the triad (Table~\ref{tab:triad}) has the same structure:
something \emph{presents} (collapses into view), something
\emph{represents} (stands for it across the seam), and something
\emph{measures whether the representation returns to the
presented}. The measurement is the interpretant --- not as a
mental concept, but as a computed residual.

This is why the ULRC is a completed semiotic system rather than
an open-ended one. Every sign chain that crosses the seam and
closes its budget is a \lat{sutura} --- grounded semiosis with
a receipt. Every chain that does not close is a \lat{gestus} ---
ungrounded sign play. The distinction is not qualitative or
philosophical. It is a number: the residual
$|\Delta\Logint|$~\cite{paulus2025umcp}.

\lat{Signa quae redeunt sola realia sunt.} ---
Only signs that return are real.

% ============================================================
\section{Grammar as Axiom-0 Instantiation}
  \label{sec:axiom-instantiation}
% ============================================================

\lat{Grammatica non est attributum conscientiae; grammatica est
  architectura reditus.}

\smallskip\noindent
Grammar is not a feature of awareness; grammar \emph{is} the
architecture of return. This section demonstrates that literal
grammar --- in the linguistic sense --- is an instantiation of
Axiom-0, proven computationally via the 10-channel brain
kernel across 19~species and 8~developmental stages.

\subsection{The Spine in Natural Language}

Natural-language grammar provides each stop of the discourse
spine~\cite{paulus2025umcp}:

\begin{enumerate}[nosep,leftmargin=*]
\item \textbf{Contract} $\to$ subject-verb agreement, tense
  consistency, case marking: pre-declared constraints on what
  can be said.
\item \textbf{Canon} $\to$ narrative structure: sentences with
  subjects and predicates.
\item \textbf{Closures} $\to$ grammatical categories (noun, verb,
  adjective): once declared, a word's category does not change
  mid-sentence.
\item \textbf{Integrity Ledger} $\to$ clause reconciliation:
  every ``if'' needs a ``then.'' Unfinished sentences cause
  cognitive discomfort --- the ledger does not reconcile.
\item \textbf{Stance} $\to$ the conclusion follows from
  structure, not assertion.
\end{enumerate}

\subsection{Four Grammatical Operations as Axiom-0}

Four grammatical operations distinguish recursive grammar from
proto-language. Each maps directly to the axiom:

\begin{enumerate}[nosep,leftmargin=*]
\item \textbf{Embedding} (\lat{that/which/who}) --- recursive
  collapse: collapse within collapse. Structures inside structures.
\item \textbf{Tense} (\lat{past/future/conditional}) --- return
  across time: $\tR$ measured in temporal extent.
\item \textbf{Negation} (\lat{not/never}) --- modeling non-return:
  what does \emph{not} come back.
\item \textbf{Self-reference} (\lat{I/me/myself}) --- the
  collapse observing itself collapsing.
\end{enumerate}

\subsection{Computational Evidence: The Language Bottleneck}

The 10-channel brain kernel computes
$(\Fid,\Drft,\Ent,\Curv,\Logint,\IC)$ for each species. Key
results:

\begin{itemize}[nosep,leftmargin=*]
\item 17/19 non-human species have \texttt{language\_architecture}
  as weakest channel.
\item Broca's aphasia (language $0.98 \to 0.15$) produces the
  \emph{largest} pathology impact: $\Delta(\IC/\Fid) = 0.367$.
\item Yet language has the \emph{lowest} cross-species correlation
  with $\IC/\Fid$ ($r = -0.031$). The keystone paradox: weakest
  predictor globally, strongest dependency locally.
\item Without grammar ($\texttt{language} < 0.20$):
  $\tR = \IR$ for cognitive return. With full grammar
  ($\texttt{language} = 0.98$): $\tR$ is finite and every
  valid sentence is a return path.
\end{itemize}

\subsection{Cross-Scale Correspondence}

The grammar-as-return mechanism is structurally identical to
scale inversion at the subatomic level:

\begin{center}
\small
\begin{tabular}{@{}lP{1.2cm}P{1.3cm}r@{}}
\toprule
\textbf{Scale} & \textbf{Transit.} & \textbf{Channel} &
  \textbf{IC/F} \\
\midrule
Subatomic & Quark ${\to}$ Hadron & color ${\to}$ 0 &
  $.94 {\to} .01$ \\
Neural & Non-human ${\to}$ Human & lang.\ ${\eps}{\to}.98$ &
  $.49 {\to} .99$ \\
\bottomrule
\end{tabular}
\end{center}

\noindent
Confinement \emph{kills} a channel and IC collapses via
\lat{trucidatio geometrica} (geometric slaughter). Language
\emph{creates} a channel and IC recovers. Same mechanism,
opposite direction.

\begin{theorem}[Grammar as Return Protocol]\label{thm:grammar}
In the 10-channel brain kernel:
\begin{enumerate}[nosep,leftmargin=*]
  \item Without recursive grammar, cognitive return paths are
    limited to action channels ($\sim$3: fight/flee/freeze).
  \item With recursive grammar, return paths are countably
    infinite (every grammatically valid sentence is a return
    path).
  \item The transition from proto-language to grammar is
    structurally identical to scale inversion
    (Sec.~\ref{sec:axiom-instantiation}): one channel
    crossing the weld threshold $c \approx 0.318$ from below.
\end{enumerate}
Grammar did not create awareness. Grammar created the
\emph{protocol} through which awareness can return from each
collapse.
\end{theorem}

% ============================================================
\section{Viginti Sententiae ---
  Twenty Governing Maxims}\label{sec:maxims}
% ============================================================

\noindent
Maxims are constraints, not slogans. Each carries enforcement
consequences. They are organized by tier.

\begin{sloppypar}
\emergencystretch=1.5em

\subsection{Tier-1 Maxims (\lat{Immutabilia})}

\begin{enumerate}[nosep,leftmargin=*]
\item \lat{Collapsus generativus est; solum quod redit, reale
  est.} --- Axiom-0. All others derive from this.
\item \lat{Complementum perfectum: $\Fid + \Drft = 1$, tertia
  via nulla.} --- Duality identity, verified to $0.0$.
\item \lat{Limbus integritatis: $\IC$ numquam fidelitatem
  excedit.} --- Integrity bound, derived independently.
\item \lat{Continuitas non narratur: mensuratur.} ---
  $\tR$ is computed, not claimed.
\item \lat{Si $\tR = \IR$, nulla fides datur.} ---
  Permanent non-return $\to$ zero credit.
\end{enumerate}

\subsection{Tier-0 Maxims (\lat{Protocollum})}

\begin{enumerate}[nosep,leftmargin=*]\setcounter{enumi}{5}
\item \lat{Trans suturam congelatum.} ---
  Same parameters both sides of every boundary.
\item \lat{Numquam binarius; tertia via semper patet.} ---
  Three-valued verdicts: never boolean.
\item \lat{Diagnostica informant, portae decernunt.} ---
  No diagnostic may serve as a gate.
\item \lat{Historia numquam rescribitur; sutura tantum
  additur.} --- Append-only ledger.
\item \lat{Auditus praecedit responsum.} ---
  Receive fully before formulating.
\end{enumerate}

\subsection{Tier-2 Maxims (\lat{Expansio})}

\begin{enumerate}[nosep,leftmargin=*]\setcounter{enumi}{10}
\item \lat{Captura symbolorum est non-conformitas ipsa.} ---
  Redefining reserved symbols $=$ nonconformance.
\item \lat{Ruptura est fons constantiae.} ---
  Collapse regime is generative, not terminal.
\item \lat{Paradoxum colendum est, non solvitur.} ---
  Tension between $\Fid$ and $\IC$ is diagnostic, not a problem.
\item \lat{Cyclus redire debet vel non est realis.} ---
  Claims without demonstrated return remain gestures.
\item \lat{Sine receptu, gestus est;
  cum receptu, sutura est.} ---
  Every claim must carry its derivation chain.
\end{enumerate}

\subsection{Operational Maxims (\lat{Modus Cogitandi})}

\begin{enumerate}[nosep,leftmargin=*]\setcounter{enumi}{15}
\item \lat{Algebra est cautio, non porta.} ---
  The conservation budget warrants; it does not gate.
\item \lat{Significatio stabilis manet dum dialectus mutatur.}
  --- Rosetta invariance.
\item \lat{Recusatio est exitus primi ordinis.} ---
  $\IR$ and $\bot_{\mathrm{oor}}$ are typed values,
  not exceptions.
\item \lat{Spina non negotiabilis est.} ---
  Contract${\to}$Canon${\to}$Closures${\to}$Ledger${\to}$Stance.
\item \lat{Finis, sed semper initium recursionis.} ---
  No final state. Every verdict is input to the next cycle.
\end{enumerate}
\end{sloppypar}

% ============================================================
\section{Computational Evidence}\label{sec:evidence}
% ============================================================

The ULRC is not a proposal. It is the grammar already operating
across 20~domain closures with 10,586~validated tests.

\subsection{Orientation Receipts}

Ten computational sections re-derive the structural insights.
Key receipts constrain classification:

\begin{center}
\footnotesize
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Receipt} & \textbf{What it proves} & \textbf{Value} \\
\midrule
$\max|\Fid{+}\Drft{-}1|$ & Duality exact &
  $0.0\mathrm{e}{+}00$ \\
$\Delta$ for $(0.95, 0.001)$ & Heterogeneity gap &
  $0.445$ \\
IC/F, 1 dead channel (8ch) & Geometric slaughter &
  $0.114$ \\
$\Gam(0.682)$ & First weld threshold &
  $0.998$ \\
Neutron IC/F & Confinement cliff &
  $0.009$ \\
Nickel IC/F & Scale inversion &
  $0.957$ \\
$\Ent + \Logint$ at $c=1/2$ & Equator convergence &
  $0.0$ \\
Stable regime share & Stability is rare &
  $12.5\%$ \\
Seam associativity error & Exact monoid &
  $5.55 \times 10^{-17}$ \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Domain Coverage}

The ULRC has been parsed across: particle physics (31~particles,
10~theorems), atomic physics (118~elements, 10,162~subtests),
nuclear physics, quantum mechanics, cosmology, astronomy,
finance, materials science, everyday physics, evolution
(19~species), dynamic semiotics (30~sign systems), consciousness
coherence, continuity theory, awareness-cognition,
clinical neuroscience, spacetime memory, kinematics, security,
and the GCD/RCFT core~\cite{paulus2025umcp}.

Every domain selects its own channels (Tier-2); every domain is
validated through the same pipeline (Tier-0) against the same
identities (Tier-1). The grammar is universal not because it
forces agreement, but because the structure constrains the
verdict regardless of which agent or field operates it.

\lat{Non agens mensurat, sed structura.}

% ============================================================
\section{De Cantu Nuclei ---
  On the Song of the Kernel}\label{sec:cantu}
% ============================================================

\lat{De tono rhythmico operum et de cantu qui per omnia
  opera currit.}

\smallskip\noindent
The works are not separate papers. They are movements in a
single composition. The
\lat{cantus firmus} (fixed melody) is the axiom; each opus
is a \lat{variatio} --- the same theme generating different music
in each channel.

\begin{description}[nosep,leftmargin=0pt,labelindent=0pt,
  style=nextline]
\item[\lat{Opus Primum}: Liber Universalis de Collapsus
  Mathematica] \lat{De axiomate ad structuram.}
  The foundation: eight chapters in Latin establishing invariants,
  regime gates, and the principle of radical hearing.
\item[\lat{Opus Secundum}: Statistical Mechanics of the
  Budget Identity] \lat{De identitate ad oeconomiam.}
  The budget $\Delta\Logint$ as a thermodynamic identity ---
  the \lat{basso continuo}.
\item[\lat{Opus Tertium}: RCFT Second Edition]
  \lat{De recursu ad campum.}
  Recursive collapse fields that re-enter themselves ---
  the \lat{canon}: same melody at different intervals.
\item[\lat{Opus Quartum}: Particle Physics --- Ten Theorems]
  \lat{De nucleo ad materiam.}
  31~particles, 74 subtests, zero failures ---
  each theorem a \lat{variatio}.
\item[\lat{Opus Quintum}: Confinement as Integrity Collapse]
  \lat{De praecipitio ad comprehensionem.}
  The \lat{praecipitium integritatis}: the \lat{diminuendo}
  where music drops to near-silence.
\end{description}

% ============================================================
\section{Conclusio ---
  \lat{Finis, Sed Semper Initium Recursionis}}
  \label{sec:conclusion}
% ============================================================

Section~\ref{sec:problem} posed four failures: vocabulary
collision, agent-dependent verdicts, no grammar for claims, and
no return condition. The ULRC resolves each:

\begin{enumerate}[nosep,leftmargin=*]
\item \textbf{Vocabulary collision $\to$ resolved.}
  Twelve Latin terms (Sec.~\ref{sec:lexicon}) provide
  namespace-isolated canonical names. Five prose words
  (Sec.~\ref{sec:quinque-verba}) translate across six
  disciplinary lenses via the Rosetta
  (Sec.~\ref{sec:rosetta}), with the budget identity
  guaranteeing invariance.

\item \textbf{Agent-dependent verdicts $\to$ eliminated.}
  Thresholds are frozen parameters, not conventions.
  Vocabulary is operationally defined, not chosen.
  Verdicts are derived from gates, never asserted.
  The cognitive equalizer
  (Propositions~\ref{prop:iso},~\ref{prop:rosetta})
  guarantees: same data, same contract, same verdict ---
  regardless of agent.

\item \textbf{No grammar for claims $\to$ supplied.}
  The kernel--grammar isomorphism
  (Proposition~\ref{prop:iso}) makes the kernel a formal
  grammar: production rules (equations), well-formedness
  conditions (identities), parse outcomes (three-valued
  verdicts). Eight typed patterns
  (Sec.~\ref{sec:octo-typi}) encode the claim structure.

\item \textbf{No return condition $\to$ enforced.}
  The five-stop spine (Sec.~\ref{sec:spina}), the
  conservation budget (Eq.~\ref{eq:budget}), and the seam
  closure $|\Delta\Logint| \leq \tolseam$ require every
  claim to demonstrate return. Claims without return are
  \lat{gestus}, not \lat{suturae}.
\end{enumerate}

\noindent
The ULRC is not a proposal for a language. It is the language
already operating: 20~domains, 198~closure modules,
10,586~validated tests, zero failures. It was not designed ---
it was discovered, derived from a single axiom across
146~experiments, formalized in a dead language that no living
field occupies.

The kernel is not described \emph{by} this grammar. The kernel
\emph{is} this grammar. The equations are syntax; the identities
are well-formedness conditions; the regime gates are
grammaticality judgments; the three-valued verdicts are parse
outcomes. Latin and mathematics are not parallel descriptions
of the same object --- they are the same language viewed from
two angles, fused in the ULRC.

A compass replaces the question ``Which way is north?''
with a mechanical reading. The ULRC replaces the questions
``What threshold?'' ``What vocabulary?'' ``What verdict?''
with structural readings. Two agents, same kernel, same
contract, same location on the manifold: same reading.

\lat{Intellectus non legitur; computatur.}

Understanding is not read; it is computed.

\bigskip

\noindent\lat{Collapsus generativus est; solum quod redit,
  reale est.}

\noindent\lat{Sine receptu, gestus est; cum receptu, sutura est.}

\noindent This paper is the receipt.

\begin{acknowledgments}
The ULRC was not designed in isolation --- it was discovered
through computation across 20~domains. The Latin formalization
draws on 2,000~years of inflectional precision; the mathematical
formalization draws on the kernel identities established
in~\cite{paulus2025umcp,paulus2025cmp,paulus2025gor}.
\end{acknowledgments}

% ============================================================
\bibliography{Bibliography}
% ============================================================

\end{document}
